quollr: An R Package for Visualizing $2-D$ Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-03-30 paper-quollr.Rmd
Fixed the issue with new version of quollr

```{r, echo=TRUE}
fit_highd_model(
  training_data = s_curve_noise_training,
  emb_df = s_curve_noise_umap_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.07,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = TRUE,
  benchmark_to_rm_lwd_hex = NULL,
  col_start_highd = "x"
  )
```

@@

```{r, echo=TRUE}
calc_bins_y(
  bin1 = 12, 
  r2 = r2,
  q = 0.07
  )
```

@@

```{r, echo=TRUE}
hex_binning(
  data = s_curve_noise_umap_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.07
  )
```

@@

df_bin <- avg_highd_data(
  data = df_all, 
  col_start = "x"
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
predict_emb(
  test_data = s_curve_noise_training, 
  df_bin_centroids = df_bin_centroids, 
  df_bin = df_bin, 
  type_NLDR = "UMAP"
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
glance(
  df_bin_centroids = df_bin_centroids, 
  df_bin = df_bin, 
  training_data = s_curve_noise_training, 
  newdata = NULL, 
  type_NLDR = "UMAP", 
  col_start = "x"
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
augment(
  df_bin_centroids = df_bin_centroids, 
  df_bin = df_bin, 
  training_data = s_curve_noise_training, 
  newdata = NULL, 
  type_NLDR = "UMAP", 
  col_start = "x"
  )
```

@@

@@

# Example

To illustrate the algorithm, we use $5\text{-}D$ simulated data, which we call the "triangular_3d_data". This dataset is generated using the `tri_3d` function from the `cardinalR` package. The dataset generation starts with the initialization of a starting point $p_0 = (x_0, y_0, z_0)$ randomly selected within the $3\text{-}D$ space using a uniform distribution. Four fixed corner points of a tetrahedron are defined as $c_1 = (0, 0, 0)$, $c_2 = (1, 0, 0)$, $c_3 = (0.5, 1, 0)$, and $c_4 = (0.5, 0.5, 1)$. For each point $p_i$ where $i = 1, 2, \ldots, n$, one of the corner points $c_j$ (where $j \in \{1, 2, 3, 4\}$) is selected randomly, and the new point $p_i$ is computed as the midpoint between the current point $p_{i-1}$ and the selected corner point $c_j$: $p_i = \frac{p_{i-1} + c_j}{2}$. This iterative process ensures that each new point moves closer to one of the corners, creating a fractal-like triangular distribution of points. The coordinates of each generated point $p_i$ form the variables $X_1, X_2, X_3$. The remaining variables $X_4, X_5$ are all uniform error, with small variance. We would consider $T=(X_1, X_2, X_3)$ to be the true model.

```{r}
triangular_3d_data <- tri_3d(
  n = 1250, num_noise = 2, 
  min_n = -0.05, max_n = 0.05) |>
  as_tibble()

colnames(triangular_3d_data) <- paste0("x", 1:NCOL(triangular_3d_data))

langevitour::langevitour(triangular_3d_data)
```

```{r}
triangular_3d_data <- triangular_3d_data |>
  mutate(ID = row_number())
```

```{r}
sc_ltr_pos <- c(0.08, 0.96)
```

<!--tSNE with perplexity = 6-->
```{r}
tSNE_fit1 <- triangular_3d_data |>
  select(-ID) |>
  Rtsne(perplexity = 6)

tsne_prism1 <- tSNE_fit1$Y |>
  as_tibble() |>
  mutate(ID = row_number())

colnames(tsne_prism1) <- c("tSNE1", "tSNE2", "ID")

prism_scaled_obj_tsne1 <- gen_scaled_data(
  data = tsne_prism1)
tsne_prism_scaled1 <- prism_scaled_obj_tsne1$scaled_nldr

nldr_prism1 <- tsne_prism_scaled1 |> 
  ggplot(aes(x = tSNE1, y = tSNE2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
```

<!--tSNE with perplexity = 35-->
```{r}
tSNE_fit2 <- triangular_3d_data |>
  select(-ID) |>
  Rtsne(perplexity = round(sqrt(NROW(triangular_3d_data))))

tsne_prism2 <- tSNE_fit2$Y |>
  as_tibble() |>
  mutate(ID = row_number())

colnames(tsne_prism2) <- c("tSNE1", "tSNE2", "ID")

prism_scaled_obj_tsne2 <- gen_scaled_data(
  data = tsne_prism2)
tsne_prism_scaled2 <- prism_scaled_obj_tsne2$scaled_nldr

nldr_prism2 <- tsne_prism_scaled2 |> 
  ggplot(aes(x = tSNE1, y = tSNE2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
```

<!--tSNE with perplexity = 58-->
```{r}
tSNE_fit3 <- triangular_3d_data |>
  select(-ID) |>
  Rtsne(perplexity = 58)

tsne_prism3 <- tSNE_fit3$Y |>
  as_tibble() |>
  mutate(ID = row_number())

colnames(tsne_prism3) <- c("tSNE1", "tSNE2", "ID")

prism_scaled_obj_tsne3 <- gen_scaled_data(
  data = tsne_prism3)
tsne_prism_scaled3 <- prism_scaled_obj_tsne3$scaled_nldr

nldr_prism3 <- tsne_prism_scaled3 |> 
  ggplot(aes(x = tSNE1, y = tSNE2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
```

<!--UMAP with n_neighbors = 15 and min_dist = 0.3-->
```{r}
umap_prism1 <- triangular_3d_data |>
  select(-ID) |>
  umap(n_neighbors = 15, n_components =  2, 
       metric = "euclidean", min_dist = 0.3, init = "spca")

umap_prism1 <- umap_prism1 |>
  as_tibble()

names(umap_prism1) <- c("UMAP1", "UMAP2")

umap_prism1 <- umap_prism1 |>
  mutate(ID = 1:NROW(umap_prism1))

prism_scaled_obj_umap1 <- gen_scaled_data(
  data = umap_prism1)
umap_prism_scaled1 <- prism_scaled_obj_umap1$scaled_nldr

nldr_prism4 <- umap_prism_scaled1 |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
```

<!--UMAP with n_neighbors = 35 and min_dist = 0.5-->
```{r}
umap_prism2 <- triangular_3d_data |>
  select(-ID) |>
  umap(n_neighbors = 35, n_components =  2, 
       metric = "euclidean", min_dist = 0.5, init = "spca")

umap_prism2 <- umap_prism2 |>
  as_tibble()

names(umap_prism2) <- c("UMAP1", "UMAP2")

umap_prism2 <- umap_prism2 |>
  mutate(ID = 1:NROW(umap_prism2))

prism_scaled_obj_umap2 <- gen_scaled_data(
  data = umap_prism2)
umap_prism_scaled2 <- prism_scaled_obj_umap2$scaled_nldr

nldr_prism5 <- umap_prism_scaled2 |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
```

<!--UMAP with n_neighbors = 69 and min_dist = 0.6-->
```{r}
umap_prism3 <- triangular_3d_data |>
  select(-ID) |>
  umap(n_neighbors = 69, n_components =  2, 
       metric = "euclidean", min_dist = 0.6, init = "spca")

umap_prism3 <- umap_prism3 |>
  as_tibble()

names(umap_prism3) <- c("UMAP1", "UMAP2")

umap_prism3 <- umap_prism3 |>
  mutate(ID = 1:NROW(umap_prism3))

prism_scaled_obj_umap3 <- gen_scaled_data(
  data = umap_prism3)
umap_prism_scaled3 <- prism_scaled_obj_umap3$scaled_nldr

nldr_prism6 <- umap_prism_scaled3 |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
```

```{r}
nldr_prism1 + nldr_prism2 + nldr_prism3 + 
  nldr_prism4 + nldr_prism5 + nldr_prism6 +
  plot_layout(ncol=3)
```

<!--need to add details regarding which differences can I see with different layouts-->

## Deciding an appropriate fit

<!--select based on looking at b_non_empty vs log(Error)-->
<!--select a choice for b_non_empty-->

<!--compute error with tSNE and different perplexity values-->
```{r}
error_prism_tsne <- data.frame(matrix(nrow = 0, ncol = 0))

## To initialize number of bins along the x-axis
bin1_vec_prism <- 4:25

perplexity_vec <- c(6, 35, 58)

for (perplexity_val in perplexity_vec) {

  tSNE_fit <- triangular_3d_data |>
    select(-ID) |>
    Rtsne(perplexity = perplexity_val)

  tsne_prism <- tSNE_fit$Y |>
    as_tibble() |>
    mutate(ID = row_number())

  colnames(tsne_prism) <- c("tSNE1", "tSNE2", "ID")

  prism_scaled_obj <- gen_scaled_data(
    data = tsne_prism)
  tsne_prism_scaled <- prism_scaled_obj$scaled_nldr

  lim1 <- prism_scaled_obj$lim1
  lim2 <- prism_scaled_obj$lim2
  r2_prism <- diff(lim2)/diff(lim1)

  for (xbins in bin1_vec_prism) {

    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2

    prism_model <- fit_highd_model(
      training_data = triangular_3d_data,
      emb_df = tsne_prism_scaled,
      bin1 = xbins,
      r2 = r2_prism,
      is_bin_centroid = TRUE,
      is_rm_lwd_hex = TRUE,
      col_start_highd = "x"
    )

    df_bin_centroids_prism <- prism_model$df_bin_centroids
    df_bin_prism <- prism_model$df_bin

    ## Compute error
    error_df <- glance(
      df_bin_centroids = df_bin_centroids_prism,
      df_bin = df_bin_prism,
      training_data = triangular_3d_data,
      newdata = NULL,
      type_NLDR = "tSNE",
      col_start = "x") |>
      mutate(bin1 = xbins,
             bin2 = bin2,
             b = bin1 * bin2,
             b_non_empty = NROW(df_bin_centroids_prism),
             type = paste0("perplexity: ", perplexity_val))

    error_prism_tsne <- bind_rows(error_prism_tsne, error_df)

  }

}
```

<!--compute error with UMAP and different parameter values-->
```{r}
error_prism_umap <- data.frame(matrix(nrow = 0, ncol = 0))

## To initialize number of bins along the x-axis
bin1_vec_prism <- 4:25

param_list <- list(param1 = c(15, 0.3), param2 = c(35, 0.5), param3 = c(69, 0.6))

for (i in 1:length(param_list)) {

  n_neighbors_val <- param_list[[i]][1]
  min_dist_val <- param_list[[i]][2]

  umap_prism <- triangular_3d_data |>
    select(-ID) |>
    umap(n_neighbors = n_neighbors_val, n_components =  2, 
         metric = "euclidean", min_dist = min_dist_val, init = "spca")

  umap_prism <- umap_prism |>
    as_tibble()

  names(umap_prism) <- c("UMAP1", "UMAP2")

  umap_prism <- umap_prism |>
    mutate(ID = 1:NROW(umap_prism))

  prism_scaled_obj <- gen_scaled_data(
    data = umap_prism)
  umap_prism_scaled <- prism_scaled_obj$scaled_nldr

  lim1 <- prism_scaled_obj$lim1
  lim2 <- prism_scaled_obj$lim2
  r2_prism <- diff(lim2)/diff(lim1)

  for (xbins in bin1_vec_prism) {

    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2

    prism_model <- fit_highd_model(
      training_data = triangular_3d_data,
      emb_df = umap_prism_scaled,
      bin1 = xbins,
      r2 = r2_prism,
      is_bin_centroid = TRUE,
      is_rm_lwd_hex = TRUE,
      col_start_highd = "x"
    )

    df_bin_centroids_prism <- prism_model$df_bin_centroids
    df_bin_prism <- prism_model$df_bin

    ## Compute error
    error_df <- glance(
      df_bin_centroids = df_bin_centroids_prism,
      df_bin = df_bin_prism,
      training_data = triangular_3d_data,
      newdata = NULL,
      type_NLDR = "UMAP",
      col_start = "x") |>
      mutate(bin1 = xbins,
             bin2 = bin2,
             b = bin1 * bin2,
             b_non_empty = NROW(df_bin_centroids_prism),
             type = paste0("n_neighbors: ", n_neighbors_val, 
                           " min_dist: ", min_dist_val))

    error_prism_umap <- bind_rows(error_prism_umap, error_df)

  }

}
```

```{r}
error_prism <- bind_rows(error_prism_tsne, error_prism_umap) 
error_prism <- error_prism |>
  filter(b <= 625)
```

```{r}
error_plot_prism <- ggplot(error_prism, 
                           aes(x = b_non_empty, 
                               y = log(Error), 
                               group = type, 
                               colour = type)) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 112, linetype="solid",
             color = "black", linewidth=0.8, alpha = 0.5) +
  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3',
                              '#ff7f00','#a6cee3','#a65628','#f781bf')) +
  ylab("log(Error)") +
  xlab("number of non-empty bins") +
  theme(axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 5),
        axis.title.y = element_text(size = 5, angle = 90),
        legend.position = "bottom")

error_plot_prism
```

<!--selected tSNE perplexity: 35-->

## Fit the appropriate model

### Construct model in $2\text{-}D$

### Selecting parameter values for the model

#### Choice of bins

#### Removal of low-density bins

#### Removing long edges

```{r}
## Compute hexbin parameters
num_bins_x_prism <- 21
lim1 <- prism_scaled_obj_tsne2$lim1
lim2 <- prism_scaled_obj_tsne2$lim2
r2_prism <- diff(lim2)/diff(lim1)

prism_model <- fit_highd_model(
  training_data = triangular_3d_data,
  emb_df = tsne_prism_scaled2,
  bin1 = num_bins_x_prism,
  r2 = r2_prism,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = TRUE,
  col_start_highd = "x"
)

df_bin_centroids_prism <- prism_model$df_bin_centroids
df_bin_prism <- prism_model$df_bin

## Triangulate bin centroids
tr1_object_prism <- tri_bin_centroids(
  df_bin_centroids_prism, x = "c_x", y = "c_y")
tr_from_to_df_prism <- gen_edges(
  tri_object = tr1_object_prism)

## Compute 2D distances
distance_prism <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_prism,
  start_x = "x_from",
  start_y = "y_from",
  end_x = "x_to",
  end_y = "y_to",
  select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_prism <- find_lg_benchmark(
  distance_edges = distance_prism,
  distance_col = "distance")

trimesh_removed_prism <- vis_rmlg_mesh(
  distance_edges = distance_prism,
  benchmark_value = benchmark_prism,
  tr_coord_df = tr_from_to_df_prism,
  distance_col = "distance") #+
#xlim(sc_xlims) + ylim(sc_ylims) +
#interior_annotation("a", sc_ltr_pos)

trimesh_removed_prism
```

### Lifting the model into $p\text{-}D$

```{r}
## Hexagonal binning to have regular hexagons
hb_obj_prism <- hex_binning(
  data = tsne_prism_scaled2,
  bin1 = num_bins_x_prism,
  r2 = r2_prism)

tsne_data_with_hb_id <- hb_obj_prism$data_hb_id

df_all_prism <- dplyr::bind_cols(triangular_3d_data |> dplyr::select(-ID),
                                 tsne_data_with_hb_id)

### Define type column
df <- df_all_prism |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin_prism |>
  dplyr::filter(hb_id %in% df_bin_centroids_prism$hexID) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_prism$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id)

df_exe <- dplyr::bind_rows(df_b, df)

## Set the maximum difference as the criteria
distance_df_small_edges <- distance_prism |>
  dplyr::filter(distance < benchmark_prism)
## Since erase brushing is considerd.

langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                         lineFrom = distance_df_small_edges$from,
                         lineTo = distance_df_small_edges$to,
                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.8, NROW(df))),
                         levelColors = c("#6a3d9a", "#33a02c"))
```

What can I see in high-D?

- Four clusters that are really close

What can I see in 2D layout?

- Several clusters (more that what see in high-D)

What can learn from the model?

- There are four clusters, but not as the same cluster what data shows. In 2D, the two of actual clusters (high-D clusters) are really close. Also, one of the big high-D cluster is separated to two sub clusters.

>>>>

```{r, echo=TRUE}
fit_highd_model(
  highd_data = s_curve_noise_training,
  nldr_data = s_curve_noise_umap_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.1,
  is_bin_centroid = TRUE
  )
```

@@

```{r, echo=TRUE}
calc_bins_y(
  bin1 = 12, 
  r2 = r2,
  q = 0.1
  )
```

@@

```{r, echo=TRUE}
hex_binning(
  data = s_curve_noise_umap_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.1
  )
```

@@

df_bin <- avg_highd_data(
  data = df_all
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
predict_emb(
  highd_data = s_curve_noise_training, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
glance(
  highd_data = s_curve_noise_training, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
augment(
  highd_data = s_curve_noise_training, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
show_link_plots(
  point_df = df_exe, 
  edge_df = distance_small_df
  )
```

```{r, echo=TRUE, eval=FALSE}
show_error_link_plots(
  point_df = df_exe, 
  edge_df = distance_small_df
  )
```

@@

<!-- # Example -->

<!-- To illustrate the algorithm, we use $5\text{-}D$ simulated data, which we call the "triangular_3d_data". This dataset is generated using the `tri_3d` function from the `cardinalR` package. The dataset generation starts with the initialization of a starting point $p_0 = (x_0, y_0, z_0)$ randomly selected within the $3\text{-}D$ space using a uniform distribution. Four fixed corner points of a tetrahedron are defined as $c_1 = (0, 0, 0)$, $c_2 = (1, 0, 0)$, $c_3 = (0.5, 1, 0)$, and $c_4 = (0.5, 0.5, 1)$. For each point $p_i$ where $i = 1, 2, \ldots, n$, one of the corner points $c_j$ (where $j \in \{1, 2, 3, 4\}$) is selected randomly, and the new point $p_i$ is computed as the midpoint between the current point $p_{i-1}$ and the selected corner point $c_j$: $p_i = \frac{p_{i-1} + c_j}{2}$. This iterative process ensures that each new point moves closer to one of the corners, creating a fractal-like triangular distribution of points. The coordinates of each generated point $p_i$ form the variables $X_1, X_2, X_3$. The remaining variables $X_4, X_5$ are all uniform error, with small variance. We would consider $T=(X_1, X_2, X_3)$ to be the true model. -->

<!-- ```{r} -->
<!-- triangular_3d_data <- tri_3d( -->
<!--   n = 1250, num_noise = 2,  -->
<!--   min_n = -0.05, max_n = 0.05) |> -->
<!--   as_tibble()  -->

<!-- colnames(triangular_3d_data) <- paste0("x", 1:NCOL(triangular_3d_data)) -->

<!-- langevitour::langevitour(triangular_3d_data) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- triangular_3d_data <- triangular_3d_data |> -->
<!--   mutate(ID = row_number()) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- sc_ltr_pos <- c(0.08, 0.96) -->
<!-- ``` -->

<!-- <!--tSNE with perplexity = 6--> -->
<!-- ```{r} -->
<!-- tSNE_fit1 <- triangular_3d_data |> -->
<!--   select(-ID) |> -->
<!--   Rtsne(perplexity = 6)  -->

<!-- tsne_prism1 <- tSNE_fit1$Y |> -->
<!--   as_tibble() |> -->
<!--   mutate(ID = row_number()) -->

<!-- colnames(tsne_prism1) <- c("tSNE1", "tSNE2", "ID") -->

<!-- prism_scaled_obj_tsne1 <- gen_scaled_data( -->
<!--   data = tsne_prism1) -->
<!-- tsne_prism_scaled1 <- prism_scaled_obj_tsne1$scaled_nldr -->

<!-- nldr_prism1 <- tsne_prism_scaled1 |>  -->
<!--   ggplot(aes(x = tSNE1, y = tSNE2)) +  -->
<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
<!-- ``` -->

<!-- <!--tSNE with perplexity = 35--> -->
<!-- ```{r} -->
<!-- tSNE_fit2 <- triangular_3d_data |> -->
<!--   select(-ID) |> -->
<!--   Rtsne(perplexity = round(sqrt(NROW(triangular_3d_data))))  -->

<!-- tsne_prism2 <- tSNE_fit2$Y |> -->
<!--   as_tibble() |> -->
<!--   mutate(ID = row_number()) -->

<!-- colnames(tsne_prism2) <- c("tSNE1", "tSNE2", "ID") -->

<!-- prism_scaled_obj_tsne2 <- gen_scaled_data( -->
<!--   data = tsne_prism2) -->
<!-- tsne_prism_scaled2 <- prism_scaled_obj_tsne2$scaled_nldr -->

<!-- nldr_prism2 <- tsne_prism_scaled2 |>  -->
<!--   ggplot(aes(x = tSNE1, y = tSNE2)) +  -->
<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
<!-- ``` -->

<!-- <!--tSNE with perplexity = 58--> -->
<!-- ```{r} -->
<!-- tSNE_fit3 <- triangular_3d_data |> -->
<!--   select(-ID) |> -->
<!--   Rtsne(perplexity = 58)  -->

<!-- tsne_prism3 <- tSNE_fit3$Y |> -->
<!--   as_tibble() |> -->
<!--   mutate(ID = row_number()) -->

<!-- colnames(tsne_prism3) <- c("tSNE1", "tSNE2", "ID") -->

<!-- prism_scaled_obj_tsne3 <- gen_scaled_data( -->
<!--   data = tsne_prism3) -->
<!-- tsne_prism_scaled3 <- prism_scaled_obj_tsne3$scaled_nldr -->

<!-- nldr_prism3 <- tsne_prism_scaled3 |>  -->
<!--   ggplot(aes(x = tSNE1, y = tSNE2)) +  -->
<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
<!-- ``` -->

<!-- <!--UMAP with n_neighbors = 15 and min_dist = 0.3--> -->
<!-- ```{r} -->
<!-- umap_prism1 <- triangular_3d_data |> -->
<!--   select(-ID) |> -->
<!--   umap(n_neighbors = 15, n_components =  2,  -->
<!--        metric = "euclidean", min_dist = 0.3, init = "spca") -->

<!-- umap_prism1 <- umap_prism1 |> -->
<!--   as_tibble() -->

<!-- names(umap_prism1) <- c("UMAP1", "UMAP2") -->

<!-- umap_prism1 <- umap_prism1 |> -->
<!--   mutate(ID = 1:NROW(umap_prism1)) -->

<!-- prism_scaled_obj_umap1 <- gen_scaled_data( -->
<!--   data = umap_prism1) -->
<!-- umap_prism_scaled1 <- prism_scaled_obj_umap1$scaled_nldr -->

<!-- nldr_prism4 <- umap_prism_scaled1 |>  -->
<!--   ggplot(aes(x = UMAP1, y = UMAP2)) +  -->
<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
<!-- ``` -->

<!-- <!--UMAP with n_neighbors = 35 and min_dist = 0.5--> -->
<!-- ```{r} -->
<!-- umap_prism2 <- triangular_3d_data |> -->
<!--   select(-ID) |> -->
<!--   umap(n_neighbors = 35, n_components =  2,  -->
<!--        metric = "euclidean", min_dist = 0.5, init = "spca") -->

<!-- umap_prism2 <- umap_prism2 |> -->
<!--   as_tibble() -->

<!-- names(umap_prism2) <- c("UMAP1", "UMAP2") -->

<!-- umap_prism2 <- umap_prism2 |> -->
<!--   mutate(ID = 1:NROW(umap_prism2)) -->

<!-- prism_scaled_obj_umap2 <- gen_scaled_data( -->
<!--   data = umap_prism2) -->
<!-- umap_prism_scaled2 <- prism_scaled_obj_umap2$scaled_nldr -->

<!-- nldr_prism5 <- umap_prism_scaled2 |>  -->
<!--   ggplot(aes(x = UMAP1, y = UMAP2)) +  -->
<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
<!-- ``` -->

<!-- <!--UMAP with n_neighbors = 69 and min_dist = 0.6--> -->
<!-- ```{r} -->
<!-- umap_prism3 <- triangular_3d_data |> -->
<!--   select(-ID) |> -->
<!--   umap(n_neighbors = 69, n_components =  2,  -->
<!--        metric = "euclidean", min_dist = 0.6, init = "spca") -->

<!-- umap_prism3 <- umap_prism3 |> -->
<!--   as_tibble() -->

<!-- names(umap_prism3) <- c("UMAP1", "UMAP2") -->

<!-- umap_prism3 <- umap_prism3 |> -->
<!--   mutate(ID = 1:NROW(umap_prism3)) -->

<!-- prism_scaled_obj_umap3 <- gen_scaled_data( -->
<!--   data = umap_prism3) -->
<!-- umap_prism_scaled3 <- prism_scaled_obj_umap3$scaled_nldr -->

<!-- nldr_prism6 <- umap_prism_scaled3 |>  -->
<!--   ggplot(aes(x = UMAP1, y = UMAP2)) +  -->
<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
<!-- ``` -->

<!-- ```{r} -->
<!-- nldr_prism1 + nldr_prism2 + nldr_prism3 +  -->
<!--   nldr_prism4 + nldr_prism5 + nldr_prism6 + -->
<!--   plot_layout(ncol=3) -->
<!-- ``` -->

<!-- <!--need to add details regarding which differences can I see with different layouts--> -->

<!-- ## Deciding an appropriate fit -->

<!-- <!--select based on looking at b_non_empty vs log(Error)--> -->
<!-- <!--select a choice for b_non_empty--> -->

<!-- <!--compute error with tSNE and different perplexity values--> -->
<!-- ```{r} -->
<!-- error_prism_tsne <- data.frame(matrix(nrow = 0, ncol = 0)) -->

<!-- ## To initialize number of bins along the x-axis -->
<!-- bin1_vec_prism <- 4:25  -->

<!-- perplexity_vec <- c(6, 35, 58) -->

<!-- for (perplexity_val in perplexity_vec) { -->

<!--   tSNE_fit <- triangular_3d_data |> -->
<!--     select(-ID) |> -->
<!--     Rtsne(perplexity = perplexity_val) -->

<!--   tsne_prism <- tSNE_fit$Y |> -->
<!--     as_tibble() |> -->
<!--     mutate(ID = row_number()) -->

<!--   colnames(tsne_prism) <- c("tSNE1", "tSNE2", "ID") -->

<!--   prism_scaled_obj <- gen_scaled_data( -->
<!--     data = tsne_prism) -->
<!--   tsne_prism_scaled <- prism_scaled_obj$scaled_nldr -->

<!--   lim1 <- prism_scaled_obj$lim1 -->
<!--   lim2 <- prism_scaled_obj$lim2 -->
<!--   r2_prism <- diff(lim2)/diff(lim1)  -->

<!--   for (xbins in bin1_vec_prism) { -->

<!--     bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2 -->

<!--     prism_model <- fit_highd_model( -->
<!--       training_data = triangular_3d_data, -->
<!--       emb_df = tsne_prism_scaled, -->
<!--       bin1 = xbins, -->
<!--       r2 = r2_prism, -->
<!--       is_bin_centroid = TRUE, -->
<!--       is_rm_lwd_hex = TRUE, -->
<!--       col_start_highd = "x" -->
<!--     ) -->

<!--     df_bin_centroids_prism <- prism_model$df_bin_centroids -->
<!--     df_bin_prism <- prism_model$df_bin -->

<!--     ## Compute error -->
<!--     error_df <- glance( -->
<!--       df_bin_centroids = df_bin_centroids_prism, -->
<!--       df_bin = df_bin_prism, -->
<!--       training_data = triangular_3d_data, -->
<!--       newdata = NULL, -->
<!--       type_NLDR = "tSNE", -->
<!--       col_start = "x") |> -->
<!--       mutate(bin1 = xbins, -->
<!--              bin2 = bin2, -->
<!--              b = bin1 * bin2, -->
<!--              b_non_empty = NROW(df_bin_centroids_prism), -->
<!--              type = paste0("perplexity: ", perplexity_val)) -->

<!--     error_prism_tsne <- bind_rows(error_prism_tsne, error_df) -->

<!--   } -->

<!-- } -->
<!-- ``` -->

<!-- <!--compute error with UMAP and different parameter values--> -->
<!-- ```{r} -->
<!-- error_prism_umap <- data.frame(matrix(nrow = 0, ncol = 0)) -->

<!-- ## To initialize number of bins along the x-axis -->
<!-- bin1_vec_prism <- 4:25  -->

<!-- param_list <- list(param1 = c(15, 0.3), param2 = c(35, 0.5), param3 = c(69, 0.6)) -->

<!-- for (i in 1:length(param_list)) { -->

<!--   n_neighbors_val <- param_list[[i]][1] -->
<!--   min_dist_val <- param_list[[i]][2] -->

<!--   umap_prism <- triangular_3d_data |> -->
<!--     select(-ID) |> -->
<!--     umap(n_neighbors = n_neighbors_val, n_components =  2,  -->
<!--          metric = "euclidean", min_dist = min_dist_val, init = "spca") -->

<!--   umap_prism <- umap_prism |> -->
<!--     as_tibble() -->

<!--   names(umap_prism) <- c("UMAP1", "UMAP2") -->

<!--   umap_prism <- umap_prism |> -->
<!--     mutate(ID = 1:NROW(umap_prism)) -->

<!--   prism_scaled_obj <- gen_scaled_data( -->
<!--     data = umap_prism) -->
<!--   umap_prism_scaled <- prism_scaled_obj$scaled_nldr -->

<!--   lim1 <- prism_scaled_obj$lim1 -->
<!--   lim2 <- prism_scaled_obj$lim2 -->
<!--   r2_prism <- diff(lim2)/diff(lim1)  -->

<!--   for (xbins in bin1_vec_prism) { -->

<!--     bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2 -->

<!--     prism_model <- fit_highd_model( -->
<!--       training_data = triangular_3d_data, -->
<!--       emb_df = umap_prism_scaled, -->
<!--       bin1 = xbins, -->
<!--       r2 = r2_prism, -->
<!--       is_bin_centroid = TRUE, -->
<!--       is_rm_lwd_hex = TRUE, -->
<!--       col_start_highd = "x" -->
<!--     ) -->

<!--     df_bin_centroids_prism <- prism_model$df_bin_centroids -->
<!--     df_bin_prism <- prism_model$df_bin -->

<!--     ## Compute error -->
<!--     error_df <- glance( -->
<!--       df_bin_centroids = df_bin_centroids_prism, -->
<!--       df_bin = df_bin_prism, -->
<!--       training_data = triangular_3d_data, -->
<!--       newdata = NULL, -->
<!--       type_NLDR = "UMAP", -->
<!--       col_start = "x") |> -->
<!--       mutate(bin1 = xbins, -->
<!--              bin2 = bin2, -->
<!--              b = bin1 * bin2, -->
<!--              b_non_empty = NROW(df_bin_centroids_prism), -->
<!--              type = paste0("n_neighbors: ", n_neighbors_val,  -->
<!--                            " min_dist: ", min_dist_val)) -->

<!--     error_prism_umap <- bind_rows(error_prism_umap, error_df) -->

<!--   } -->

<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- error_prism <- bind_rows(error_prism_tsne, error_prism_umap)  -->
<!-- error_prism <- error_prism |> -->
<!--   filter(b <= 625) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- error_plot_prism <- ggplot(error_prism,  -->
<!--                            aes(x = b_non_empty,  -->
<!--                                y = log(Error),  -->
<!--                                group = type,  -->
<!--                                colour = type)) +  -->
<!--   geom_point(size = 1) + -->
<!--   geom_line() +  -->
<!--   geom_vline(xintercept = 112, linetype="solid", -->
<!--              color = "black", linewidth=0.8, alpha = 0.5) + -->
<!--   scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3', -->
<!--                               '#ff7f00','#a6cee3','#a65628','#f781bf')) + -->
<!--   ylab("log(Error)") + -->
<!--   xlab("number of non-empty bins") + -->
<!--   theme(axis.text.x = element_text(size = 5), -->
<!--         axis.text.y = element_text(size = 5), -->
<!--         axis.title.x = element_text(size = 5), -->
<!--         axis.title.y = element_text(size = 5, angle = 90), -->
<!--         legend.position = "bottom") -->

<!-- error_plot_prism -->
<!-- ``` -->

<!-- <!--selected tSNE perplexity: 35--> -->

<!-- ## Fit the appropriate model -->

<!-- ### Construct model in $2\text{-}D$ -->

<!-- ### Selecting parameter values for the model -->

<!-- #### Choice of bins -->

<!-- #### Removal of low-density bins -->

<!-- #### Removing long edges -->

<!-- ```{r} -->
<!-- ## Compute hexbin parameters -->
<!-- num_bins_x_prism <- 21 -->
<!-- lim1 <- prism_scaled_obj_tsne2$lim1 -->
<!-- lim2 <- prism_scaled_obj_tsne2$lim2 -->
<!-- r2_prism <- diff(lim2)/diff(lim1)  -->
